---
title: "Data Visualization With ggplot2"
author: "Adejumo Ridwan Suleiman"
date: "`r Sys.Date()`"
output: html_document
---

# Installing and Loading ggplot2
```{r}
#install.packages("ggplot2")
library(ggplot2)
```

# Diamonds dataset
```{r}
head(diamonds)
```
# Economics dataset
```{r}
head(economics)
```

# Barplots

## Vertical Bar Plot
```{r}
# Most occurring diamonds (change fill with "fill = ", change line colour with "colour = ")
diamonds |> 
  ggplot(mapping = aes(x = cut)) +
  geom_bar()
```

## Horizontal Bar Plot
```{r}
diamonds |> 
  ggplot(mapping = aes(y = cut)) +
  geom_bar()
```

## Stacked BarPlot
```{r}
diamonds |> 
  ggplot(mapping = aes(x = cut, fill = clarity), position = "fill") +
  geom_bar()
```
## Grouped Barplot
```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = clarity), position = "dodge")
```

# Boxplots

## Comparing diamond cut and price
```{r}
diamonds |> 
  ggplot(mapping = aes(x = cut, y = price, colour = clarity)) +
  geom_boxplot()
```

# Histogram
```{r}
#Check for the distribution of diamond carat
diamonds |> 
  ggplot(aes(x = carat)) +
  geom_histogram()
```

# Scatter Plot
```{r}
# Check for the relationship between diamond price and carat
diamonds |> 
  ggplot(aes(x = price, y = carat)) +
  geom_point()
```

# Line Graph
```{r}
economics |> 
  ggplot(aes(x = date, y = pop)) +
  geom_line()
```


# Labelling
```{r}
diamonds |> 
  ggplot(mapping = aes(x = cut, y = price, colour = cut)) +
  geom_boxplot() +
  labs(x = "Cut", y = "Price", title = "Diamond Cut against Price", )
```

# Facetting
```{r}
diamonds |> 
  ggplot(mapping = aes(y = cut, fill = cut)) +
  geom_bar() +
  facet_grid(color ~ .)
```


#Exercise

## Question 1
Using a boxplot to visualize diamond cut against depth percentage. Which diamond cut have the highest median depth percentage.
1. fair
2. Good
3. Very Good
4. Premium
5. Ideal
```{r}
diamonds |> 
  ggplot(mapping = aes(x = cut, y = depth, color = cut)) +
  geom_boxplot()
```
## Question 2
Visualize using a line graph the change in United States Personal consumption expenditures in the economics data. In which year did the United States have the low personal consumption expenditure.
1. 1967
2. 2000
3. 1999
4. 2015
5. 1990
```{r}
economics |> 
  ggplot(mapping = aes(x = date, y = pce)) +
  geom_line()
```
## Question 3
Using a suitable bin width, visualize the distribution of diamond carat in the diamonds data set. In which range of carat value do most diamonds fall in.
1. 0.5 - 1.5
2. 4.0 - 5.0
3. 2.0 - 3.5
4. 0.0 - 0.5
```{r}
diamonds |> 
  ggplot(mapping = aes(x = carat)) +
  geom_histogram()
```
# Question 4
Which of the folling can I use to plot a categorical variable against a categorical variable
1. Boxplot
2. Scatterplot
3. Histogram
4. LinePlot
5. None of the Above

# Question 5
Which of the following codes in the options, will give the graph below.
```{r}
diamonds |> 
  ggplot(mapping = aes(x = cut)) +
  geom_bar()
```

```{r}
diamonds |> 
  ggplot() +
  geom_bar(mapping = aes(x = cut))
```

```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut))
```

```{r}
ggplot(data = diamonds, mapping = aes(x = cut)) +
  geom_bar()
```


